On the uniqueness and stability of dictionaries \\ for sparse representation of noisy signals

### 2019-05-10 manuscripts/IEEE/final_submission.tex
merge

Here, we prove very generally that uniqueness and stability is a typical property of sparse dictionary learning. More specifically,  we show that dictionaries injective on a sparse domain are identifiable from \mbox{$N = m(k-1){m \choose k} + m$} noisy k-sparse linear combinations of their $m$ columns up to an error that is linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $n \geq \min(2k,m)$, in almost all cases the problem is well-posed, as per Hadamard \cite{Hadamard1902}, given a sufficient amount of data (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}).

@@

Another interesting question raised by our work is for which special cases is it efficient to check that a solution to Prob.~\ref{InverseProblem} or \ref{OptimizationProblem} is unique up to noise and inherent ambiguities. Considering that the sufficient conditions we have described for checking this in general are NP-hard to compute, are the necessary conditions hard? Are Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} then also hard \cite{tilman?}? Finally, since solving Prob.~\ref{SLCopt} directly is intractable in general, but efficiently solvable by $\ell_1$-norm minimization for fixed matrices with large enough lower bounds over sparse domains \cite{?}, is there a version of Thm.~\ref{SLCopt} certifying when Prob.~\ref{OptimizationProblem} can be solved efficiently in a similar way?  % [*** Incorporate this open problem with more words ***]

>>>>

Here, we prove very generally that uniqueness and stability is a typical property of sparse dictionary learning. More specifically,  we show that dictionaries injective on a sparse domain are identifiable from \mbox{$N = m(k-1){m \choose k} + m$} noisy sparse linear combinations of $k$ of their $m$ columns up to an error that is linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $n \geq \min(2k,m)$, in almost all cases the problem is well-posed, as per Hadamard \cite{Hadamard1902}, given a sufficient amount of data (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}).

@@

Another interesting question raised by our work is for which special cases is it efficient to check that a solution to Prob.~\ref{InverseProblem} or \ref{OptimizationProblem} is unique up to noise and inherent ambiguities. Considering that the sufficient conditions we have described for checking this in general are NP-hard to compute, are the necessary conditions hard? Are Probs.~\ref{InverseProblem} and \ref{OptimizationProblem} then also hard \cite{Tillmann15}? Finally, since solving Prob.~\ref{SLCopt} directly is intractable in general, but efficiently solvable by $\ell_1$-norm minimization for fixed matrices with large enough lower bounds over sparse domains \cite{eldar2012compressed}, is there a version of Thm.~\ref{SLCopt} certifying when Prob.~\ref{OptimizationProblem} can be solved efficiently in a similar way?  % [*** Incorporate this open problem with more words ***]
